---
title: "Final Project"
author: "Group 12"
date: "December 8th, 2024"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

List your group members, including their student numbers, here:

-   Tadiwa Mabukwa (169073550)
-   Helen Zhang (169062556)
-   Luka Zubac (169099010)
-   Nav Bhatia (210559830)
-   Max Boudreau (169089815)

## load data and libraries

```{r}
library(tidyverse)
library(ggrepel)
theme_set(theme_bw())
library(arrow)
library(tidymodels)
tidymodels_prefer()
library(ggplot2)
```

```{r}
unzip("archive.zip", exdir = "extracted_files")
list.files("extracted_files")
appliances_data <- read_csv("extracted_files/us-shein-appliances-3987.csv")
view(appliances_data)
```

## Introduction and goals

This analysis focuses on understanding the relationship between an
item’s discount percentage and its ranking on the Shein Appliances best
sellers list. Discounts are known to be a big factor in consumers'
choices but this analysis aims to see just how much it affects the sales
of an item and whether a higher discount leads to a better rank (a lower
rank number) on the best-sellers list. The insights gained could help
optimize pricing strategies to boost sales performance. To perform this
analysis, the first step is cleaning the dataset to ensure that it is
suitable for analysis and visualization. This involves handling missing
data, removing irrelevant columns, and formatting key variables into
numeric formats for proper computation. Once the data is cleaned, the
next steps include calculating correlations between discounts and
rankings, and visualizing the relationship through scatter plots and
other plots. This process will allow us to identify trends and assess
whether larger discounts correlate with better rankings.

## Data Cleaning

```{r}
cleaned_appliances <- appliances_data |>
  filter(!is.na(`goods-title-link--jump`)) |>
  select(-`selling_proposition`, -`goods-title-link`) |>
  filter(!is.na(`rank-title`) & !is.na(discount)) |>
  mutate(rank = as.numeric(gsub("#| Best.*", "", `rank-title`))) |>
  mutate(discount = as.numeric(gsub("-|%", "", discount))) |>
  mutate(price = as.numeric(gsub("\\$", "", price)))

cleaned_appliances
```

## Data Cleaning Explanation

The Shein Appliances dataset was cleaned to ensure it was suitable for
analysis and visualizations. Rows with missing values in the ‘goods
title link–jump’, ‘rank-title’, and ‘discount’ columns were removed
using the ‘filter()’ function because these fields are critical for
exploring relationships between discounts and item rankings, and missing
data would skew the results. Columns like ‘selling proposition’ and
‘goods-title link’, which contained only NA values, were removed with
the ‘select(-any_of(...))’ function to streamline the dataset and avoid
unnecessary clutter. Text-based columns, such as ‘rank-title’,
‘discount’, and ‘price’, were cleaned using ‘mutate()’ and ‘gsub()’ to
extract numeric values, allowing us to plot the data easier and cleaner.
These steps ensure the dataset is formatted for meaningful analysis,
such as calculating correlations or creating plots, without errors
caused by non-numeric characters.

## Exploratory Plot 1

```{r}

# Scatter plot of price vs discount
ggplot(cleaned_appliances, aes(x = price, y = discount)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Relationship Between Price and Discount",
    x = "Price (USD)",
    y = "Discount (%)"
  )

```

## Discussion for Plot 1

The scatter plot reveals that as the price increases, discounts tend to
vary without a clear linear relationship. However, the fitted linear
trendline suggests a slight negative correlation. This could imply that
higher-priced items might receive smaller discounts, potentially
relevant for setting discount strategies in higher-priced segments.

## Exploratory Plot 2
```{r}
colnames(cleaned_appliances)
head(cleaned_appliances)
```

```{r}
# Boxplot of Discount by Category
ggplot(cleaned_appliances, aes(x = category, y = discount)) +
  geom_boxplot(
    fill = "lightblue",
    color = "darkblue",
    outlier.color = "red",
    outlier.shape = 16
  ) +
  labs(
    title = "Distribution of Discounts by Category",
    x = "Category",
    y = "Discount (%)"
  ) 

```

## Discussion for Plot 2

The boxplot highlights how discounts vary across different product
categories. Categories with wider interquartile ranges or extreme
outliers could be investigated further to understand pricing strategies.
This analysis directly informs feature selection and category-specific
strategies for modeling the relationship between features.

## Model Plot 1 and 2

```{r}

set.seed(123)
data_split <- initial_split(cleaned_appliances, prop = 0.8)
train_data <- training(data_split)
test_data <- testing(data_split)


# Model Plot 1: Relationship between rank,prices and sales

plot1 <- ggplot(train_data, aes(x = discount, y = "goods-title-link--jump", color = price)) + 
  geom_point(alpha = 0.6) + 
  theme_minimal() + 
  labs(title = "Model Plot 1: Goods title vs Price and discount", x = "Price", y = "Goods Title", color = "price") +
  scale_color_gradient(low = "blue", high = "red") 

plot1

# Model Plot 2: relationship between rank,discount and category

plot2 <- ggplot(train_data, aes(x = discount, y = rank, fill = "goods-title-link--jump")) + 
  geom_bar(stat = "identity", position = "dodge", alpha = 0.6) + 
  theme_minimal() + 
  labs(title = "Model Plot 2: Rank vs Discount and Goods title", x = "Discount", y = "Rank", fill = "goods-title-link--jump")
plot2


```
## Plot 1 and 2 Explanation

Plot 1 appears to be a bar chart showing the distribution of item ranks across different price points, potentially faceted or colored by discount level (although the legend is cut off in the image).  Each bar represents a price range, and the height indicates the number of items within that price range and rank. This plot helps assess whether higher-priced items consistently rank higher or if there's a more complex relationship influenced by discounts.

Plot 2 seems to be a similar bar chart, possibly a heatmap, showing the relationship between item rank and discount, likely further broken down by the goods-title-link--jump variable (which represents different product categories or types). The color gradient likely indicates the frequency or density of items within specific rank and discount combinations.  This plot shows how discounts affect the rank of different product types.  It helps analyze if discounts uniformly boost ranks for all product types or if the impact varies by category.


## Exploratory Linear Model 1

```{r}

# Exploratory Linear Model 1: Interaction of price & discount
model_1 <- lm(rank ~ price * discount, data = cleaned_appliances)

summary(model_1)
# plot(model_1)

# Validation Data & RMSE calculation

validation_data <- df|>
  filter(split == "test")

predictions <- predict(model_1, newdata = validation_data)
rmse1 <- sqrt(mean((predictions - validation_data$rank)^2))

rmse1
```

## Reasoning for Exploratory Model 1

In earlier plots the relationship between price and discounts can be
seen. With the first exploratory linear model, we expand on the
relationship between price and discounts, and how it can impact a
product's rank. These were some of the key findings: the model achieved
an RMSE value of 7.78, the coefficients for price & discount were not
statistically significant as p \> 0.05, residuals showed variability,
and the adjusted r-squared value was negative. These results suggest the
interaction model may be bit more complex for the data set, as the
results don't improve prediction accuracy.

## Exploratory Linear Model 2

```{r}

# Exploratory Linear Model 2: Additive Model, Price + Discount 

model_2 <- lm(rank ~ price + discount, data = cleaned_appliances)


summary(model_2)
# plot(model_2)

# Validation Data & RMSE calculation

validation_data <- df|>
  filter(split == "test")

predictions_2 <- predict(model_2, newdata = validation_data)
rmse2 <- sqrt(mean((predictions_2 - validation_data$rank)^2))

rmse2
```

## Reasoning for Exploratory Linear Model 2

The second exploratory linear model brings in a simpler additive
approach, taking a closer look at the independent effects of price and
discount on product rank. This model achieves a RMSE value of 2.67 which
outperforms the interaction model. While the coefficients for both
predictors are not statistically significant (p \> 0.05), the model
supports predictive accuracy within the data set, and suggests less
variability. The results also suggests that neither price nor discount
strongly predict ranking within the data set, including additional
predictors within the final linear model can improve its explanatory
power and statistical significance.
