---
output:
  pdf_document: default
  html_document: default
---
title: "Final Project"
author: "Group 12"
date: "December 8th, 2024"
output: pdf_document
---
List your group members, including their student numbers, here:

- Helen Zhang (169062556)
- Luka Zubac (169099010)
- Nav Bhatia (210559830)
- Tadiwa Mabukwa (169073550)

## Load data and Libraries

```{r}
library(tidyverse)
library(ggrepel)
theme_set(theme_bw())
library(arrow)
library(tidymodels)
tidymodels_prefer()
```

```{r}
unzip("archive.zip", exdir = "extracted_files")
list.files("extracted_files")
appliances_data <- read_csv("extracted_files/us-shein-appliances-3987.csv")
view(appliances_data)
```

## Introduction and Goals
This analysis focuses on understanding the relationship between an item’s discount percentage and its ranking on the Shein Appliances best
sellers list. Discounts are known to be a big factor in consumers' choices but this analysis aims to see just how much it affects the sales of
an item and whether a higher discount leads to a better rank (a lower rank number) on the best-sellers list. The insights gained could help
optimize pricing strategies to boost sales performance. To perform this analysis, the first step is cleaning the dataset to ensure that it is
suitable for analysis and visualization. This involves handling missing data, removing irrelevant columns, and formatting key variables into
numeric formats for proper computation. Once the data is cleaned, the next steps include calculating correlations between discounts and
rankings, and visualizing the relationship through scatter plots and other plots. This process will allow us to identify trends and assess
whether larger discounts correlate with better rankings.

## Data Cleaning

```{r cars}
cleaned_appliances <- appliances_data |>
  filter(!is.na(`goods-title-link--jump`)) |>
  select(-`selling_proposition`, -`goods-title-link`) |>
  filter(!is.na(`rank-title`) & !is.na(discount)) |>
  mutate(rank = as.numeric(gsub("#| Best.*", "", `rank-title`))) |>
  mutate(discount = as.numeric(gsub("-|%", "", discount))) |>
  mutate(price = as.numeric(gsub("\\$", "", price)))
cleaned_appliances
```
The Shein Appliances dataset was cleaned to ensure it was suitable for analysis and visualizations. Rows with missing values in the ‘goods
title link–jump’, ‘rank-title’, and ‘discount’ columns were removed using the ‘filter()’ function because these fields are critical for exploring
relationships between discounts and item rankings, and missing data would skew the results. Columns like ‘selling proposition’ and ‘goods-title
link’, which contained only NA values, were removed with the ‘select(-any_of(...))’ function to streamline the dataset and avoid unnecessary
clutter. Text-based columns, such as ‘rank-title’, ‘discount’, and ‘price’, were cleaned using ‘mutate()’ and ‘gsub()’ to extract numeric
values, allowing us to plot the data easier and cleaner. These steps ensure the dataset is formatted for meaningful analysis, such as
calculating correlations or creating plots, without errors caused by non-numeric characters.

## Exploratory Plot 1

```{r}
# Scatter plot of price vs discount
ggplot(cleaned_appliances, aes(x = price, y = discount)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Relationship Between Price and Discount",
    x = "Price (USD)",
    y = "Discount (%)"
  )
```
## Discussion for Plot 1
The scatter plot reveals that as the price increases, discounts tend to vary without a clear linear relationship. However, the fitted linear trendline suggests a slight negative correlation. This could imply that higher-priced items might receive smaller discounts, potentially relevant for setting discount strategies in higher-priced segments.
## Exploratory Plot 2
```{r}

# Boxplot of Discount by Category
ggplot(cleaned_appliances, aes(x ="goods-title-link--jump", y = discount)) +
  geom_boxplot(
    fill = "lightblue",
    color = "darkblue",
    outlier.color = "red",
    outlier.shape = 16
  ) +
  labs(
    title = "Distribution of Discounts by Category",
    x = "Category",
    y = "Discount (%)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
## Discussion for Plot 2
The boxplot highlights how discounts vary across different product categories. Categories with wider interquartile ranges or extreme
outliers could be investigated further to understand pricing strategies. This analysis directly informs feature selection and category-specific strategies for modeling the relationship between features.
## Model Plot 1 and 2
```{r}
set.seed(123)
data_split <- initial_split(cleaned_appliances, prop = 0.8)
train_data <- training(data_split)
test_data <- testing(data_split)

# Model Plot 1: Relationship between rank,prices and sales
plot1 <- ggplot(train_data, aes(x = discount, y = "goods-title-link--jump", color = price)) + 
  geom_point(alpha = 0.6) + 
  theme_minimal() + 
  labs(title = "Model Plot 1: Goods title vs Price and discount", x = "Price", y = "Goods Title", color = "price") +
  scale_color_gradient(low = "blue", high = "red")  
plot1

# Model Plot 2: relationship between rank,discount and category
plot2 <- ggplot(train_data, aes(x = discount, y = rank, fill = "goods-title-link--jump")) + 
  geom_bar(stat = "identity", position = "dodge", alpha = 0.6) + 
  theme_minimal() + 
  labs(title = "Model Plot 2: Rank vs Discount and Goods title", x = "Discount", y = "Rank", fill = "goods-title-link--jump")
plot2

```
## Exploratory Linear Model 1
```{r}
# Exploratory Linear Model 1: Interaction of price & discount
model_1 <- lm(rank ~ price * discount, data = cleaned_appliances)
summary(model_1)
# plot(model_1)
# Validation Data & RMSE calculation
cleaned_appliances$split <- sample(c("train", "test"), size = nrow(cleaned_appliances), replace = TRUE, prob = c(0.8, 0.2))
validation_data <- cleaned_appliances[cleaned_appliances$split == "test", ]
predictions <- predict(model_1, newdata = validation_data)
rmse1 <- sqrt(mean((predictions - validation_data$rank)^2))
rmse1
```
## Reasoning for Exploratory Model 1
In earlier plots the relationship between price and discounts can be seen. With the first exploratory linear model, we expand on the
relationship between price and discounts, and how it can impact a product's rank. These were some of the key findings: the model achieved an RMSE value of .7388399, the coefficients for price & discount were not statistically significant as p \> 0.05, residuals showed variability, and the adjusted r-squared value was negative. These results suggest the interaction model may be bit more complex for the data set, as the results don't improve prediction accuracy.
## Exploratory Linear Model 2
```{r}
# Exploratory Linear Model 2: Additive Model, Price + Discount 
model_2 <- lm(rank ~ price + discount, data = cleaned_appliances)
summary(model_2)
# plot(model_2)
# Validation Data & RMSE calculation
cleaned_appliances$split <- sample(c("train", "test"), size = nrow(cleaned_appliances), replace = TRUE, prob = c(0.8, 0.2))
validation_data <- cleaned_appliances[cleaned_appliances$split == "test", ]
predictions_2 <- predict(model_2, newdata = validation_data)
rmse2 <- sqrt(mean((predictions_2 - validation_data$rank)^2))
rmse2
```

## Reasoning for Exploratory Linear Model 2

The second exploratory linear model brings in a simpler additive approach, taking a closer look at the independent effects of price and
discount on product rank. This model achieves an RMSE value of .6780388 which outperforms the interaction model. While the coefficients for both predictors are not statistically significant (p \> 0.05), the model supports predictive accuracy within the data set and suggests less variability. The results also suggest that neither price nor discount strongly predict ranking within the data set, including additional predictors within the final linear model can improve its explanatory power and statistical significance.

## Final Linear Model Diagnostic

```{r}

# Final Linear Model

train_validation_data <- cleaned_appliances |>
  filter(!is.na(rank))

cleaned_appliances$split <- sample(c("train", "test"), size = nrow(cleaned_appliances), replace = TRUE, prob = c(0.8, 0.2))

train_validation_data <- cleaned_appliances[cleaned_appliances$split == "train", ]
test_data <- cleaned_appliances[cleaned_appliances$split == "test", ]

final_model <- lm(rank ~ price + discount, data = train_validation_data)
final_test_predictions <- predict(final_model, newdata = test_data)
final_rmse <- sqrt(mean((final_test_predictions - test_data$rank)^2, na.rm = TRUE))

cat("Final Model Summary:\n")
summary(final_model)

cat("\nTest Set RMSE:\n")
final_rmse

# Exploring relationship through different model

library(randomForest)

rf_model <- randomForest(rank ~ price + discount, data = train_validation_data)
rf_test_predictions <- predict(rf_model, newdata = test_data)
rf_rmse <- sqrt(mean((rf_test_predictions - test_data$rank)^2, na.rm = TRUE))

cat("RF Model Summary:\n")
summary(rf_model)

cat("Random Forest Test Set RMSE: ")
rf_rmse
#summary(final_rmse)
```

## Final Linear Model Interpretation

The final linear model output closely mirrors the results from the
earlier exploratory linear analysis. This similarity strengthens our
hypothesis suggesting that our initial understanding aligns with the
model's findings. Alternatively for a more precise and advanced
interpretation, it would be beneficial to explore alternative models
such as decision tree and the random forest model shown above, which
help in capture non-linear relationships, that may not be fully
represented in a linear model.

## Conclusion

The analysis aimed to discuss whether the discount percentage affects
the ranking on Shein’s appliance best seller’s list. While the initial
exploratory model indicates some potential influences between ranking
and discount percentage, the final linear model shows the goods’
categories may play a more significant role in rank. Despite this,
neither price nor discount alone has a strong, statistically significant
impact on rank. This suggests that while discounts may have some effect,
other factors, such as product category, customer reviews, or seasonal
trends, likely have a more substantial influence on product rankings.
Optimizing pricing and discount strategies may still be useful, but a
more comprehensive approach that considers multiple factors would likely
provide a more effective model for predicting sales performance.

## Limitations

The analysis has several limitations that should be considered when
interpreting the result. One major limitation is the dataset’s focus on
price, discount and product category, which leaves out other potential
variables that could significantly influence ranking, such as
customers’s reviews, brand recognition and stock levels. This
simplification of the model to only include price and discount as
predictor may overlook more complex interactions within the data.
Furthermore, the absence of temporal data ( such as seasonality or
promotional events) limits the models’ ability to account for time-based
trends, which could significantly affect product ranking. Lastly, the
discount and rank may not be as strong as initially hypothesized, and
additional data and variables may be necessary for more accurate
predictions.

## References

Measurable AI. (2024, September 20). The psychology of discounts: Temu
vs Shein vs Shopee vs Lazada vs TikTok Shop. Measurable.ai.
<https://blog.measurable.ai/2024/09/20/the-psychology-of-discounts-temu-vs-shein-vs-shopee-vs-lazada-tiktok-shop/>

The article discusses the psychological impact of discounts on consumer
behavior, comparing platforms like Shein, which relates to your research
on how discounts affect product rankings. Understanding this
relationship can help refine strategies for optimizing pricing and
boosting sales performance.

eCommerceDB. (2024, September 20). Shein business strategy, growth, and
competition. Retrieved from
<https://ecommercedb.com/insights/shein-business-strategy-growth-competition/4708>

This article examines Shein's growth strategy and competition, which can
offer valuable insights into how pricing and discount strategies
influence brand success, potentially affecting product rankings.
Understanding Shein's business approach can help optimize pricing models
in your analysis of discounts and rankings.

TwoOrbits. (2024, September 15). Marketing strategy of Shein exposed.
Retrieved from
<https://twoorbits.com/digital-marketing/marketing-strategy-of-shein-exposed/>

This article highlights Shein's marketing tactics, including pricing and
discount strategies, which are critical for understanding how these
factors can influence product rankings. By examining Shein's approach,
your analysis can gain insights into the broader competitive landscape
and how discounting affects consumer purchasing decisions, potentially
impacting product rank.

CBRE. (2023, October 20). Omnichannel: How does e-commerce vary by
product category? Retrieved from
<https://www.cbre.com/insights/articles/omnichannel-how-does-e-commerce-vary-by-product-category>

This article explores how e-commerce strategies differ across product
categories, which can help in understanding how discounts might affect
rankings on Shein's appliance category. It provides valuable insights
into how consumer behavior and category-specific marketing strategies
can influence sales and ranking outcomes in your analysis.

Zhang, Y. (2023). The rise of SHEIN: Navigating the digital era of fast
fashion and its comprehensive impacts. ResearchGate.
<https://www.researchgate.net/publication/379918500_The_Rise_of_SHEIN_Navigating_the_Digital_Era_of_Fast_Fashion_and_Its_Comprehensive_Impacts>

This article provides an analysis of SHEIN's success in the digital era,
focusing on its marketing strategies, fast-fashion model, and global
reach. It offers insights into how SHEIN has become a dominant player in
e-commerce, which can help explain the factors influencing product
rankings on its platform, including pricing and discount strategies. The
findings can support your analysis of how discounts might affect Shein's
appliance category rankings.

Sari, R. A., & Asmara, D. T. (2020). The effect of discounts on consumer
buying interest in e-commerce: Study on consumers who buy discounted
products on e-commerce Tokopedia. International Journal of Advanced
Engineering and Management, 3(3), 1-9.
<https://ijaem.net/issue_dcp/The%20Effect%20of%20Discounts%20on%20Consumer%20Buying%20Interest%20in%20E%20Commerce;%20Study%20on%20consumers%20who%20buy%20discounted%20products%20on%20E%20commerce%20Tokopedia.pdf>


