---
title: "Final Project"
author: "Group 12"
date: "December 8th, 2024"
output: pdf_document
---
List your group members, including their student numbers, here:

- 
- 
- Luka Zubac (169099010)
- Nav (210559830)
- Max Boudreau (169089815)

## load data and libraries

```{r}
library(tidyverse)
library(ggrepel)
theme_set(theme_bw())
library(arrow)
library(tidymodels)
tidymodels_prefer()
```

```{r}
unzip("archive.zip", exdir = "extracted_files")
list.files("extracted_files")
appliances_data <- read_csv("extracted_files/us-shein-appliances-3987.csv")
view(appliances_data)
```

## Introduction and goals
This analysis focuses on understanding the relationship between an item’s discount percentage and its ranking on the Shein Appliances best
sellers list. Discounts are known to be a big factor in consumers' choices but this analysis aims to see just how much it affects the sales of
an item and whether a higher discount leads to a better rank (a lower rank number) on the best-sellers list. The insights gained could help
optimize pricing strategies to boost sales performance. To perform this analysis, the first step is cleaning the dataset to ensure that it is
suitable for analysis and visualization. This involves handling missing data, removing irrelevant columns, and formatting key variables into
numeric formats for proper computation. Once the data is cleaned, the next steps include calculating correlations between discounts and
rankings, and visualizing the relationship through scatter plots and other plots. This process will allow us to identify trends and assess
whether larger discounts correlate with better rankings.

## Data Cleaning

```{r cars}
cleaned_appliances <- appliances_data |>
  filter(!is.na(`goods-title-link--jump`)) |>
  select(-`selling_proposition`, -`goods-title-link`) |>
  filter(!is.na(`rank-title`) & !is.na(discount)) |>
  mutate(rank = as.numeric(gsub("#| Best.*", "", `rank-title`))) |>
  mutate(discount = as.numeric(gsub("-|%", "", discount))) |>
  mutate(price = as.numeric(gsub("\\$", "", price)))

cleaned_appliances
```
The Shein Appliances dataset was cleaned to ensure it was suitable for analysis and visualizations. Rows with missing values in the ‘goods
title link–jump’, ‘rank-title’, and ‘discount’ columns were removed using the ‘filter()’ function because these fields are critical for exploring
relationships between discounts and item rankings, and missing data would skew the results. Columns like ‘selling proposition’ and ‘goods-title
link’, which contained only NA values, were removed with the ‘select(-any_of(...))’ function to streamline the dataset and avoid unnecessary
clutter. Text-based columns, such as ‘rank-title’, ‘discount’, and ‘price’, were cleaned using ‘mutate()’ and ‘gsub()’ to extract numeric
values, allowing us to plot the data easier and cleaner. These steps ensure the dataset is formatted for meaningful analysis, such as
calculating correlations or creating plots, without errors caused by non-numeric characters.










