---
title: "Final Project"
author: "Group 12"
date: "December 8th, 2024"
output: pdf_document
---
List your group members, including their student numbers, here:

- 
- 
- Luka Zubac (169099010)
- Nav (210559830)
- Max Boudreau (169089815)

## load data and libraries

```{r}
library(tidyverse)
library(ggrepel)
theme_set(theme_bw())
library(arrow)
library(tidymodels)
tidymodels_prefer()
```

```{r}
unzip("archive.zip", exdir = "extracted_files")
list.files("extracted_files")
appliances_data <- read_csv("extracted_files/us-shein-appliances-3987.csv")
view(appliances_data)
```

## Introduction and goals
This analysis focuses on understanding the relationship between an item’s discount percentage and its ranking on the Shein Appliances best
sellers list. Discounts are known to be a big factor in consumers' choices but this analysis aims to see just how much it affects the sales of
an item and whether a higher discount leads to a better rank (a lower rank number) on the best-sellers list. The insights gained could help
optimize pricing strategies to boost sales performance. To perform this analysis, the first step is cleaning the dataset to ensure that it is
suitable for analysis and visualization. This involves handling missing data, removing irrelevant columns, and formatting key variables into
numeric formats for proper computation. Once the data is cleaned, the next steps include calculating correlations between discounts and
rankings, and visualizing the relationship through scatter plots and other plots. This process will allow us to identify trends and assess
whether larger discounts correlate with better rankings.

## Data Cleaning

```{r cars}
cleaned_appliances <- appliances_data |>
  filter(!is.na(`goods-title-link--jump`)) |>
  select(-`selling_proposition`, -`goods-title-link`) |>
  filter(!is.na(`rank-title`) & !is.na(discount)) |>
  mutate(rank = as.numeric(gsub("#| Best.*", "", `rank-title`))) |>
  mutate(discount = as.numeric(gsub("-|%", "", discount))) |>
  mutate(price = as.numeric(gsub("\\$", "", price)))

cleaned_appliances
```
The Shein Appliances dataset was cleaned to ensure it was suitable for analysis and visualizations. Rows with missing values in the ‘goods
title link–jump’, ‘rank-title’, and ‘discount’ columns were removed using the ‘filter()’ function because these fields are critical for exploring
relationships between discounts and item rankings, and missing data would skew the results. Columns like ‘selling proposition’ and ‘goods-title
link’, which contained only NA values, were removed with the ‘select(-any_of(...))’ function to streamline the dataset and avoid unnecessary
clutter. Text-based columns, such as ‘rank-title’, ‘discount’, and ‘price’, were cleaned using ‘mutate()’ and ‘gsub()’ to extract numeric
values, allowing us to plot the data easier and cleaner. These steps ensure the dataset is formatted for meaningful analysis, such as
calculating correlations or creating plots, without errors caused by non-numeric characters.

## Exploratory Plot 1

```{r}

# Scatter plot of price vs discount
ggplot(cleaned_appliances, aes(x = price, y = discount)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Relationship Between Price and Discount",
    x = "Price (USD)",
    y = "Discount (%)"
  )

```

## Discussion for Plot 1

The scatter plot reveals that as the price increases, discounts tend to vary without a clear linear relationship. However, the fitted linear trendline suggests a slight negative correlation. This could imply that higher-priced items might receive smaller discounts, potentially relevant for setting discount strategies in higher-priced segments.

## Exploratory Plot 2

```{r}
# Boxplot of Discount by Category
ggplot(cleaned_appliances, aes(x = category, y = discount)) +
  geom_boxplot(fill = "lightblue", color = "darkblue", outlier.color = "red", outlier.shape = 16) +
  labs(
    title = "Distribution of Discounts by Category",
    x = "Category",
    y = "Discount (%)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Discussion for Plot 2

The boxplot highlights how discounts vary across different product categories. Categories with wider interquartile ranges or extreme
outliers could be investigated further to understand pricing strategies. This analysis directly informs feature selection and category-specific strategies for modeling the relationship between features.

## Model Plot 1 and 2

```{r}

set.seed(123)
df <- tibble(
split = c("train", "train", "test", "train", "test"),
category = c("A", "B", "A", "C", "B"),
price = c(10, 20, 15, 25, 30),
sales = c(100, 150, 120, 130, 140),
discount = c(0.1, 0.2, 0.15, 0.3, 0.25),
rank = c(3, 1, 2, 4, 5)  )
train_data <- df %>% filter(split == "train")

# Model Plot 1: Relationship between rank,prices and sales

plot1 <- ggplot(train_data, aes(x = price, y = sales, color = rank)) + 
geom_point(alpha = 0.6) + 
theme_minimal() + 
labs(title = "Model Plot 1: Rank vs Price and Sales", x = "Price", y = "Sales", color = "Rank") +
scale_color_gradient(low = "blue", high = "red")  

plot1

# Model Plot 2: relationship between rank,discount and category

plot2 <- ggplot(train_data, aes(x = discount, y = rank, fill = category)) + 
geom_bar(stat = "identity", position = "dodge", alpha = 0.6) + 
theme_minimal() + 
labs(title = "Model Plot 2: Rank vs Discount and Category", x = "Discount", y = "Rank", fill = "Category")

plot2


```

## Exploratory Linear Model 1

```{r}

# Exploratory Linear Model 1: Interaction of price & discount
model_1 <- lm(rank ~ price * discount, data = cleaned_appliances)

summary(model_1)
# plot(model_1)

# Validation Data & RMSE calculation

validation_data <- df|>
  filter(split == "test")

predictions <- predict(model_1, newdata = validation_data)
rmse1 <- sqrt(mean((predictions - validation_data$rank)^2))

rmse1
```

## Reasoning for Exploratory Model 1

In earlier plots the relationship between price and discounts can be seen. With the first exploratory linear model, we expand on the
relationship between price and discounts, and how it can impact a product's rank. These were some of the key findings: the model achieved an RMSE value of 7.78, the coefficients for price & discount were not statistically significant as p \> 0.05, residuals showed variability, and the adjusted r-squared value was negative. These results suggest the interaction model may be bit more complex for the data set, as the results don't improve prediction accuracy.

## Exploratory Linear Model 2

```{r}

# Exploratory Linear Model 2: Additive Model, Price + Discount 

model_2 <- lm(rank ~ price + discount, data = cleaned_appliances)


summary(model_2)
# plot(model_2)

# Validation Data & RMSE calculation

validation_data <- df|>
  filter(split == "test")

predictions_2 <- predict(model_2, newdata = validation_data)
rmse2 <- sqrt(mean((predictions_2 - validation_data$rank)^2))

rmse2
```

## Reasoning for Exploratory Linear Model 2

The second exploratory linear model brings in a simpler additive approach, taking a closer look at the independent effects of price and
discount on product rank. This model achieves an RMSE value of 2.67 which outperforms the interaction model. While the coefficients for both predictors are not statistically significant (p \> 0.05), the model supports predictive accuracy within the data set and suggests less variability. The results also suggest that neither price nor discount strongly predict ranking within the data set, including additional predictors within the final linear model can improve its explanatory power and statistical significance.






